---
title: "MetaFlowTrain: a highly-parallelized and modular fluidic system to study exometabolite-mediated inter-organismal communication"
author: "Chesneau et al., 2025, Guillaume Chesneau"
date: "06/09/2024"
output:
  html_document:
    toc: yes
    toc_float: yes
  word_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=FALSE, autodep = TRUE, warning = FALSE)
#Set working direction
knitr::opts_knit$set(root.dir = "") 
```

This script has been made on R 4.2.2 (and Rstudio 2022.12.0+353)
```{r R version}
R.version
#Load theme set
library(ggplot2)
theme_set(theme_bw())
```



```{r, echo=FALSE, warning=FALSE, message=FALSE, results="hide"}
#Palette color

#Prepare palette color for conditions

  name_ORDGANISM <- c("Media.Bacteria", "Media.Fungi", "Bacteria.Fungi", "Fungi.Bacteria", "Media.BacteriaFungi", "Media.Media")
  
  # desired color palette
  Palette_ORGANISM <- c("#ff3300", "#00cc44", "#ffd9b3", "#b35900", "#4d0026","#cccccc")
  
  # associated EnvType level
  names(Palette_ORGANISM) <- name_ORDGANISM
  
  
#Prepare palette color for conditions 2

  name_ORDGANISM2 <- c("Bacteria", "Fungi", "Bacteria_Fungi", "Fungi_Bacteria", "Bacteria_Fungi_Media", "Media_Media")
  
  # desired color palette
  Palette_ORGANISM2 <- c("#ff3300", "#00cc44", "#ffd9b3", "#b35900", "#4d0026","#cccccc")
  
  # associated EnvType level
  names(Palette_ORGANISM2) <- name_ORDGANISM2
  
  
  #Prepare palette color for conditions 2

  name_ORDGANISM3 <- c("Bacteria-Media", "Fungi-Media", "Bacteria-Fungi", "Fungi-Bacteria", "Bacteria+Fungi-Media", "Media-Media")
  
  # desired color palette
  Palette_ORGANISM3 <- c("#ff3300", "#00cc44", "#ffd9b3", "#b35900", "#4d0026","#cccccc")
  
  # associated EnvType level
  names(Palette_ORGANISM3) <- name_ORDGANISM3
  
  #Prepare palette color for TREATMENT_1
  name_TREATMENT_1 <- c("Peat", "Peat.A", "Peat.A.B", "Peat.A.BF")
  
  # desired color palette
  Palette_TREATMENT_1 <- c("#ff3300", "#00cc44", "#ffd9b3", "#b35900")
  
  # associated EnvType level
  names(Palette_TREATMENT_1) <- name_TREATMENT_1
  
str(Palette_TREATMENT_1)


  #Prepare palette color for TREATMENT_1
  name_TREATMENT_2 <- c("Peat", "Peat-Plant", "Peat-Plant-B", "Peat-Plant-B-F")
  # desired color palette
  Palette_TREATMENT_2 <- c("#a07437", "#849146", "#d13505", "#2caaaa")
  
  # associated EnvType level
  names(Palette_TREATMENT_2) <- name_TREATMENT_2
```


*****

# 1 -  Preprocessing of data

Input data are raw peak area table with samples in rows and metabolites in columns.
Data are preprocessed individually for each run and scaled in order to be compared between runs.


```{r, echo=FALSE, warning=FALSE, message=FALSE, results="hide", eval="FALSE"}
#- a) Amino acids

#---(i) Prepare files

#Open table
# This input file is a full table with all data from the study
 #https://www.intechopen.com/chapters/52527

#AA
Met <- read.table(file="Raw_AA.txt",sep="\t",dec = ".",header=TRUE)
Met
``` 


```{r, echo=FALSE, warning=FALSE, message=FALSE, results="hide", eval="FALSE"}
#---(ii) Process filtering-normalization
#Data are already TIC normalized

library(tidyr)
library(dplyr)

met_TIC <- Met #Change with any subset files to process

#Remove outliers defined by a loading difference of greater than 90% (70% in the paper) compared to the average of the total abundance sums in the run ((https://link.springer.com/article/10.1007/s11306-019-1564-8#Sec2))
# If TIC is greater then 90% I removed

met_TIC$Filtered<-ifelse(met_TIC$CORRECTION2 > 1.95,"OUT",
                         ifelse(met_TIC$CORRECTION2 < 0.05,"OUT", "IN"))

#Check number of samples removes
table(met_TIC$Filtered) #Loose 70 samples (19.49% of samples)


#calculate CV (Coefficient of Variation), allow to know metabolites that are too variable. Not really robust. NOT DO
keep <- colnames(met_TIC)[18:37]
Met_AA <- tidyr::gather(met_TIC, key =  Metabolites, value = PEAK_AREA_TIC, keep)
Met_AA$Class = c("Amino_Acids")
Met_AA
Met_AA$PEAK_AREA_TIC <- as.numeric(Met_AA$PEAK_AREA_TIC)

Met_AAf <- Met_AA


#combine year and month into one column
Met_AAf$all_bio_cat_factor
Met_AAf2 <- aggregate(PEAK_AREA_TIC ~ ORGANISM2 + Metabolites, data = Met_AAf, FUN = sum)
Met_AAf3 <- aggregate(PEAK_AREA_TIC ~ ORGANISM2 + Metabolites, data = Met_AAf, FUN = sd)
colnames(Met_AAf3) <- c("A", "B", "sd")

Met_AAf2$Sd <- c(Met_AAf3$sd)

Met_AAf2$CV <-  Met_AAf2$Sd / Met_AAf2$PEAK_AREA_TIC * 100

#INterpreation of CV
#A CV of 0.5 means the standard deviation is half as large as the mean.
#A CV of 1 means the standard deviation is equal to the mean.
#A CV of 1.5 means the standard deviation is 1.5 times larger than the mean.

#Plot CVs

Met_AAf2$col <- c("unique")
Met_AAf2[is.na(Met_AAf2)] <- 0

library(ggplot2)
Plot_CV <- ggplot(data=Met_AAf2, 
                             aes_string(x='CV',y='col')) + 
  geom_boxplot(outlier.shape = TRUE) + 
  geom_jitter(alpha=0.5) +
  ylab ("Samples") + xlab ("CVs") +
  ggtitle("distributions of CVs") + 
  geom_vline(data = Met_AAf2, aes(xintercept = as.numeric(40)))


#REmove samples when to high CV. Should be 25% according to https://www.ncbi.nlm.nih.gov/pmc/articles/PMC6570933/.

Remove <- Met_AAf2[which(Met_AAf2$CV >40),]
Remove <- Remove$all_bio_cat_factor
#This metabolites concentration in each samples are outliers, but they will be saw in plots. Can continued normalization with file called Met_AA



#___(iii) Normalized data

#1- TIC normalization (to remove technical variation due to the period of acquisition of the scans, normalisation between samples)
#2- Log transformation (correct for heteroscedasticity, to convert multiplicative relations into additive relations, and to make skewed distributions (more) symmetric )
#3- Scaling and Centering of the data (use a dispersion measure for scaling, For pareto square root is used)

#ref : https://bmcgenomics.biomedcentral.com/articles/10.1186/1471-2164-7-142


#The noise-filtered data of individual samples were  normalized by the total intensity of all metabolites via TIC we can also calculate the non TIC data
#raw value non TIC normalized

Met_AA$CORRECTION2 <- as.numeric(Met_AA$CORRECTION2)
Met_AA$PEAK_AREA <- Met_AA$PEAK_AREA_TIC * Met_AA$CORRECTION2

#https://www.ncbi.nlm.nih.gov/pmc/articles/PMC6616221/ TO be continued


#Log transformation
Met_AA$PEAK_AREA_TIC_Log <- log10(Met_AA$PEAK_AREA_TIC)
Met_AA$PEAK_AREA_TIC_Log[Met_AA$PEAK_AREA_TIC_Log == '-Inf'] <- 0

Met_AA$PEAK_AREA_Log <- log10(Met_AA$PEAK_AREA)
Met_AA$PEAK_AREA_Log[Met_AA$PEAK_AREA_Log == '-Inf'] <- 0


#Extract Log transformed data without scaling and centering
#write.csv(Met_AA, "Final/Met_AA.csv")
```


```{r, echo=FALSE, warning=FALSE, message=FALSE, results="hide", eval="FALSE"}
#- b) TCA - Glycolysis 

#---(i) Prepare files
#Open table
# This input file is a full table with all data from the study
 #https://www.intechopen.com/chapters/52527

#AA
Met <- read.table(file="Raw_TCA.txt",sep="\t",dec = ".",header=TRUE)
Met
```





```{r, echo=FALSE, warning=FALSE, message=FALSE, results="hide", eval="FALSE"}

#---(ii) Process filtering-normalization

#___(ii) Filter data
#Data are already TIC normalized

library(tidyr)
library(dplyr)

met_TIC <- Met #Change with any subset files to process

#Remove outliers defined by a loading difference of greater than 90% (70% in the paper) compared to the average of the total abundance sums in the run ((https://link.springer.com/article/10.1007/s11306-019-1564-8#Sec2))
# If TIC is greater then 90% I removed

met_TIC$Filtered<-ifelse(met_TIC$CORRECTION2 > 1.95,"OUT",
                         ifelse(met_TIC$CORRECTION2 < 0.05,"OUT", "IN"))

#Check number of samples removes
table(met_TIC$Filtered) #Loose 43 samples (11.62% of samples)


#calculate CV (Coefficient of Variation), allow to know metabolites that are too variable. Not really robust. NOT DO
keep <- colnames(met_TIC)[18:33]
MET_TCA <- tidyr::gather(met_TIC, key =  Metabolites, value = PEAK_AREA_TIC, keep)
MET_TCA$Class = c("TCA_Glycolysis")
MET_TCA
MET_TCA$PEAK_AREA_TIC <- as.numeric(MET_TCA$PEAK_AREA_TIC)

MET_TCAf <- MET_TCA


#combine year and month into one column
MET_TCAf$all_bio_cat_factor
MET_TCAf2 <- aggregate(PEAK_AREA_TIC ~ ORGANISM2 + Metabolites, data = MET_TCAf, FUN = sum)
MET_TCAf3 <- aggregate(PEAK_AREA_TIC ~ ORGANISM2 + Metabolites, data = MET_TCAf, FUN = sd)
colnames(MET_TCAf3) <- c("A", "B", "sd")

MET_TCAf2$Sd <- c(MET_TCAf3$sd)

MET_TCAf2$CV <-  MET_TCAf2$Sd / MET_TCAf2$PEAK_AREA_TIC * 100

#INterpreation of CV
#A CV of 0.5 means the standard deviation is half as large as the mean.
#A CV of 1 means the standard deviation is equal to the mean.
#A CV of 1.5 means the standard deviation is 1.5 times larger than the mean.

#Plot CVs

MET_TCAf2$col <- c("unique")
MET_TCAf2[is.na(MET_TCAf2)] <- 0

library(ggplot2)
Plot_CV <- ggplot(data=MET_TCAf2, 
                             aes_string(x='CV',y='col')) + 
  geom_boxplot(outlier.shape = TRUE) + 
  geom_jitter(alpha=0.5) +
  ylab ("Samples") + xlab ("CVs") +
  ggtitle("distributions of CVs") + 
  geom_vline(data = MET_TCAf2, aes(xintercept = as.numeric(40)))


mean(MET_TCAf2$CV)


#REmove samples when to high CV. Should be 25% according to https://www.ncbi.nlm.nih.gov/pmc/articles/PMC6570933/.

Remove <- MET_TCAf2[which(MET_TCAf2$CV >40),]
Remove <- Remove$all_bio_cat_factor
#This metabolites concentration in each samples are outliers, but they will be saw in plots. Can continued normalization with file called MET_TCA



#___(iii) Normalized data

#1- TIC normalization (to remove technical variation due to the period of acquisition of the scans, normalisation between samples)
#2- Log transformation (correct for heteroscedasticity, to convert multiplicative relations into additive relations, and to make skewed distributions (more) symmetric )
#3- Scaling and Centering of the data (use a dispersion measure for scaling, For pareto square root is used)

#ref : https://bmcgenomics.biomedcentral.com/articles/10.1186/1471-2164-7-142


#The noise-filtered data of individual samples were  normalized by the total intensity of all metabolites via TIC we can also calculate the non TIC data
#raw value non TIC normalized

MET_TCA$CORRECTION2 <- as.numeric(MET_TCA$CORRECTION2)
MET_TCA$PEAK_AREA <- MET_TCA$PEAK_AREA_TIC * MET_TCA$CORRECTION2

#https://www.ncbi.nlm.nih.gov/pmc/articles/PMC6616221/ TO be continued


#Log transformation
MET_TCA$PEAK_AREA_TIC_Log <- log10(MET_TCA$PEAK_AREA_TIC)
MET_TCA$PEAK_AREA_TIC_Log[MET_TCA$PEAK_AREA_TIC_Log == '-Inf'] <- 0

MET_TCA$PEAK_AREA_Log <- log10(MET_TCA$PEAK_AREA)
MET_TCA$PEAK_AREA_Log[MET_TCA$PEAK_AREA_Log == '-Inf'] <- 0


#Extract Log transformed data without scaling and centering
#write.csv(MET_TCA, "Final/Met_TCA.csv")
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results="hide", eval="FALSE"}
#- c) Pareto scaling 

#(Reduce the relative importance of large values, but keep data structure partially intact), Centering : focus on the differences and not the similarities in the data (does not have to be heterocedastic, log transofmred before processing)

#---(i) AA
#Open files
Met_AA <- read.csv("Final/Met_AA.csv", header=TRUE, sep = ",",  dec=".", row.names=1)

#AA
#Parato Scaling on TIC Log_trasnformed data

Met_AA_pareto <- subset(Met_AA, select = c(BIO_SAMPLE_ID3, Metabolites, PEAK_AREA_TIC_Log))
Met_AA_pareto <- tidyr::spread(Met_AA_pareto, key =  Metabolites, value = PEAK_AREA_TIC_Log)

library(IMIFA)

#Change NA to 0
Met_AA_pareto[is.na(Met_AA_pareto)] <- 0 

#Pareto calculation
Met_AA_paretof <- data.frame(pareto_scale(Met_AA_pareto[2:21],
             centering = TRUE))
Met_AA_paretof$BIO_SAMPLE_ID3 <- Met_AA_pareto$BIO_SAMPLE_ID3

#Final Table
keep <- colnames(Met_AA_paretof)[1:20]
Met_AA_Par_LogTIC <- tidyr::gather(Met_AA_paretof, key =  Metabolites, value = TIC_Log_Scaled, keep)



#Parato Scaling on Non-TIC Log_trasnformed data

Met_AA_pareto <- subset(Met_AA, select = c(BIO_SAMPLE_ID3, Metabolites, PEAK_AREA_Log))
Met_AA_pareto <- tidyr::spread(Met_AA_pareto, key =  Metabolites, value = PEAK_AREA_Log)

library(IMIFA)

#Change NA to 0
Met_AA_pareto[is.na(Met_AA_pareto)] <- 0 

#Pareto calculation
Met_AA_paretof <- data.frame(pareto_scale(Met_AA_pareto[2:21],
             centering = TRUE))
Met_AA_paretof$BIO_SAMPLE_ID3 <- Met_AA_pareto$BIO_SAMPLE_ID3

#Final Table
keep <- colnames(Met_AA_paretof)[1:20]
Met_AA_Par_Log <- tidyr::gather(Met_AA_paretof, key =  Metabolites, value = Log_Scaled, keep)




#Parato Scaling on TIC data

Met_AA_pareto <- subset(Met_AA, select = c(BIO_SAMPLE_ID3, Metabolites, PEAK_AREA_TIC))
Met_AA_pareto <- tidyr::spread(Met_AA_pareto, key =  Metabolites, value = PEAK_AREA_TIC)

library(IMIFA)

#Change NA to 0
Met_AA_pareto[is.na(Met_AA_pareto)] <- 0 

#Pareto calculation
Met_AA_paretof <- data.frame(pareto_scale(Met_AA_pareto[2:21],
             centering = TRUE))
Met_AA_paretof$BIO_SAMPLE_ID3 <- Met_AA_pareto$BIO_SAMPLE_ID3

#Final Table
keep <- colnames(Met_AA_paretof)[1:20]
Met_AA_Par_TIC <- tidyr::gather(Met_AA_paretof, key =  Metabolites, value = TIC_Scaled, keep)




#Parato Scaling on non-TIC data

Met_AA_pareto <- subset(Met_AA, select = c(BIO_SAMPLE_ID3, Metabolites, PEAK_AREA))
Met_AA_pareto <- tidyr::spread(Met_AA_pareto, key =  Metabolites, value = PEAK_AREA)

library(IMIFA)

#Change NA to 0
Met_AA_pareto[is.na(Met_AA_pareto)] <- 0 

#Pareto calculation
Met_AA_paretof <- data.frame(pareto_scale(Met_AA_pareto[2:21],
             centering = TRUE))
Met_AA_paretof$BIO_SAMPLE_ID3 <- Met_AA_pareto$BIO_SAMPLE_ID3

#Final Table
keep <- colnames(Met_AA_paretof)[1:20]
Met_AA_Par <- tidyr::gather(Met_AA_paretof, key =  Metabolites, value = Scaled, keep)
```


```{r, echo=FALSE, warning=FALSE, message=FALSE, results="hide", eval="FALSE"}
#---(ii) TCA Glycolysis

#(Reduce the relative importance of large values, but keep data structure partially intact), Centering : focus on the differences and not the similarities in the data (does not have to be heterocedastic, log transofmred before processing)

#Open files
Met_TCA <- read.csv("Final/Met_TCA.csv", header=TRUE, sep = ",",  dec=".", row.names=1)

#TCA
#Parato Scaling on TIC Log_trasnformed data

Met_TCA_pareto <- subset(Met_TCA, select = c(BIO_SAMPLE_ID3, Metabolites, PEAK_AREA_TIC_Log))
Met_TCA_pareto <- tidyr::spread(Met_TCA_pareto, key =  Metabolites, value = PEAK_AREA_TIC_Log)

library(IMIFA)

#Change NA to 0
Met_TCA_pareto[is.na(Met_TCA_pareto)] <- 0 

#Pareto calculation
Met_TCA_paretof <- data.frame(pareto_scale(Met_TCA_pareto[2:17],
             centering = TRUE))
Met_TCA_paretof$BIO_SAMPLE_ID3 <- Met_TCA_pareto$BIO_SAMPLE_ID3

#Final Table
keep <- colnames(Met_TCA_paretof)[1:16]
Met_TCA_Par_LogTIC <- tidyr::gather(Met_TCA_paretof, key =  Metabolites, value = TIC_Log_Scaled, keep)


#Parato Scaling on Non-TIC Log_trasnformed data

Met_TCA_pareto <- subset(Met_TCA, select = c(BIO_SAMPLE_ID3, Metabolites, PEAK_AREA_Log))
Met_TCA_pareto <- tidyr::spread(Met_TCA_pareto, key =  Metabolites, value = PEAK_AREA_Log)

library(IMIFA)

#Change NA to 0
Met_TCA_pareto[is.na(Met_TCA_pareto)] <- 0 

#Pareto calculation
Met_TCA_paretof <- data.frame(pareto_scale(Met_TCA_pareto[2:17],
             centering = TRUE))
Met_TCA_paretof$BIO_SAMPLE_ID3 <- Met_TCA_pareto$BIO_SAMPLE_ID3

#Final Table
keep <- colnames(Met_TCA_paretof)[1:16]
Met_TCA_Par_Log <- tidyr::gather(Met_TCA_paretof, key =  Metabolites, value = Log_Scaled, keep)




#Parato Scaling on TIC data

Met_TCA_pareto <- subset(Met_TCA, select = c(BIO_SAMPLE_ID3, Metabolites, PEAK_AREA_TIC))
Met_TCA_pareto <- tidyr::spread(Met_TCA_pareto, key =  Metabolites, value = PEAK_AREA_TIC)

library(IMIFA)

#Change NA to 0
Met_TCA_pareto[is.na(Met_TCA_pareto)] <- 0 

#Pareto calculation
Met_TCA_paretof <- data.frame(pareto_scale(Met_TCA_pareto[2:17],
             centering = TRUE))
Met_TCA_paretof$BIO_SAMPLE_ID3 <- Met_TCA_pareto$BIO_SAMPLE_ID3

#Final Table
keep <- colnames(Met_TCA_paretof)[1:16]
Met_TCA_Par_TIC <- tidyr::gather(Met_TCA_paretof, key =  Metabolites, value = TIC_Scaled, keep)



#Parato Scaling on non-TIC data

Met_TCA_pareto <- subset(Met_TCA, select = c(BIO_SAMPLE_ID3, Metabolites, PEAK_AREA))
Met_TCA_pareto <- tidyr::spread(Met_TCA_pareto, key =  Metabolites, value = PEAK_AREA)

library(IMIFA)

#Change NA to 0
Met_TCA_pareto[is.na(Met_TCA_pareto)] <- 0 

#Pareto calculation
Met_TCA_paretof <- data.frame(pareto_scale(Met_TCA_pareto[2:17],
             centering = TRUE))
Met_TCA_paretof$BIO_SAMPLE_ID3 <- Met_TCA_pareto$BIO_SAMPLE_ID3

#Final Table
keep <- colnames(Met_TCA_paretof)[1:16]
Met_TCA_Par <- tidyr::gather(Met_TCA_paretof, key =  Metabolites, value = Scaled, keep)
```



```{r, echo=FALSE, warning=FALSE, message=FALSE, results="hide", eval="FALSE"}
#- d) Merge Data

#---(i) AA

#Open files
Met_AA <- read.csv("Final/Met_AA.csv", header=TRUE, sep = ",",  dec=".", row.names=1)
Met_AA$BIO_SAMPLE_ID3_Met <- paste( Met_AA$BIO_SAMPLE_ID3, "/", Met_AA$Metabolites)

Met_TCA <- read.csv("Final/Met_TCA.csv", header=TRUE, sep = ",",  dec=".", row.names=1)

colnames(Met_AA)
colnames(Met_AA_Par_Log)
colnames(Met_AA_Par_TIC)
colnames(Met_AA_pareto)

Met_AA_Par_LogTIC$BIO_SAMPLE_ID3_Met <- paste( Met_AA_Par_LogTIC$BIO_SAMPLE_ID3, "/", Met_AA_Par_LogTIC$Metabolites)
Met_AA_Par_LogTIC$BIO_SAMPLE_ID3 <- NULL
Met_AA_Par_LogTIC$Metabolites <- NULL
Met_AA_Par_Log$BIO_SAMPLE_ID3_Met <- paste( Met_AA_Par_Log$BIO_SAMPLE_ID3, "/", Met_AA_Par_Log$Metabolites)
Met_AA_Par_Log$BIO_SAMPLE_ID3 <- NULL
Met_AA_Par_Log$Metabolites <- NULL
Met_AA_Par_TIC$BIO_SAMPLE_ID3_Met <- paste( Met_AA_Par_TIC$BIO_SAMPLE_ID3, "/", Met_AA_Par_TIC$Metabolites)
Met_AA_Par_TIC$BIO_SAMPLE_ID3 <- NULL
Met_AA_Par_TIC$Metabolites <- NULL
Met_AA_Par$BIO_SAMPLE_ID3_Met <- paste( Met_AA_Par$BIO_SAMPLE_ID3, "/", Met_AA_Par$Metabolites)
Met_AA_Par$BIO_SAMPLE_ID3 <- NULL
Met_AA_Par$Metabolites <- NULL

Met_AA_pareto <-  merge(Met_AA, Met_AA_Par_LogTIC,  by = c("BIO_SAMPLE_ID3_Met"))  
Met_AA_pareto <-  merge(Met_AA_pareto, Met_AA_Par_Log,  by = c("BIO_SAMPLE_ID3_Met"))  
Met_AA_pareto <-  merge(Met_AA_pareto, Met_AA_Par_TIC,  by = c("BIO_SAMPLE_ID3_Met"))  
Met_AA_pareto <-  merge(Met_AA_pareto, Met_AA_Par,  by = c("BIO_SAMPLE_ID3_Met"))  
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results="hide", eval="FALSE"}
#---(ii) TCA

#Open files
Met_TCA <- read.csv("Final/Met_TCA.csv", header=TRUE, sep = ",",  dec=".", row.names=1)
Met_TCA$BIO_SAMPLE_ID3_Met <- paste( Met_TCA$BIO_SAMPLE_ID3, "/", Met_TCA$Metabolites)


Met_TCA
colnames(Met_TCA)
colnames(Met_TCA_Par_Log)
colnames(Met_TCA_Par_TIC)
colnames(Met_TCA_pareto)

Met_TCA_Par_LogTIC$BIO_SAMPLE_ID3_Met <- paste( Met_TCA_Par_LogTIC$BIO_SAMPLE_ID3, "/", Met_TCA_Par_LogTIC$Metabolites)
Met_TCA_Par_LogTIC$BIO_SAMPLE_ID3 <- NULL
Met_TCA_Par_LogTIC$Metabolites <- NULL
Met_TCA_Par_Log$BIO_SAMPLE_ID3_Met <- paste( Met_TCA_Par_Log$BIO_SAMPLE_ID3, "/", Met_TCA_Par_Log$Metabolites)
Met_TCA_Par_Log$BIO_SAMPLE_ID3 <- NULL
Met_TCA_Par_Log$Metabolites <- NULL
Met_TCA_Par_TIC$BIO_SAMPLE_ID3_Met <- paste( Met_TCA_Par_TIC$BIO_SAMPLE_ID3, "/", Met_TCA_Par_TIC$Metabolites)
Met_TCA_Par_TIC$BIO_SAMPLE_ID3 <- NULL
Met_TCA_Par_TIC$Metabolites <- NULL
Met_TCA_Par$BIO_SAMPLE_ID3_Met <- paste( Met_TCA_Par$BIO_SAMPLE_ID3, "/", Met_TCA_Par$Metabolites)
Met_TCA_Par$BIO_SAMPLE_ID3 <- NULL
Met_TCA_Par$Metabolites <- NULL

Met_TCA_pareto <-  merge(Met_TCA, Met_TCA_Par_LogTIC,  by = c("BIO_SAMPLE_ID3_Met"))  
Met_TCA_pareto <-  merge(Met_TCA_pareto, Met_TCA_Par_Log,  by = c("BIO_SAMPLE_ID3_Met"))  
Met_TCA_pareto <-  merge(Met_TCA_pareto, Met_TCA_Par_TIC,  by = c("BIO_SAMPLE_ID3_Met"))  
Met_TCA_pareto <-  merge(Met_TCA_pareto, Met_TCA_Par,  by = c("BIO_SAMPLE_ID3_Met")) 
```


```{r, echo=FALSE, warning=FALSE, message=FALSE, results="hide", eval="FALSE"}
#---(iii) MErge Final file
Met_tot_f <- rbind(Met_TCA_pareto, Met_AA_pareto)

Met_tot_f[is.na(Met_tot_f)] <- 0

#write.csv(Met_tot_f, "Final/Met_tot_f.csv")
```

The post-preprocessing result comprises a singular table featuring eight distinct columns, each housing diverse values :        

* PEAK_AREA.  
* PEARK_AREA_TIC.  
* PEAK_AREA_Log. 
* PEARK_AREA_TIC_Log. 
* PEAK_AREA_Scaled. 
* PEARK_AREA_TIC_Scaled. 
* PEAK_AREA_Log_Scaled. 
* PEARK_AREA_TIC_Log_Scaled. 

*****


# 2 - Start of the script


```{r, echo=FALSE, warning=FALSE, message=FALSE, results="hide"}
#Open dataframe
Met_Final <- read.csv("Final/Met_tot_f.csv", header=TRUE, sep = ",",  dec=".", row.names=1)


#Change 3PG 2PG

Met_Final$Metabolites <- ifelse(Met_Final$Metabolites == "X2.phosphoglyceric.acid..2PG..and..3PG.", "2PG.and.3PG",Met_Final$Metabolites) 

AA <- subset(Met_Final, !Class %in% c("TCA_Glycolysis"))
TCA <- subset(Met_Final, Class %in% c("TCA_Glycolysis"))


```


## 2.1 - Calculate PERMANOVA

Calculate differences of metabolic profiles between different Peat extract for BF-SynCom

```{r, echo=FALSE, warning=FALSE, message=FALSE, results="show"}
#PERMANOVA
library(vegan)
Working_file <- Met_Final
unique(Working_file$Metabolites)
Working_file <- subset(Working_file, ORGANISM %in% c( "X.BF", "X.X"))
Working_file <- subset(Working_file, !TREATMENT2 %in% c("Peat.A.B"))
Working_file <- subset(Working_file, !BIO_SAMPLE_ID3 %in% c("45258_48_X.X_62h_Peat_B.1", "45268_58_X.BF_62h_Peat.A.BF_B.1", "45268_59_X.X_62h_Peat.A.BF_B.1","45258_48_X.X_62h_Peat_B.4", "45258_47_X.X_62h_Peat_B.4"))
#Working_file <- subset(Working_file, Time_point %in% c("62h"))

Working_file_tot <- Working_file
library(dplyr)
Working_file_tot <- Working_file_tot %>% select(c("BIO_SAMPLE_ID3","PEAK_AREA_TIC_Log","Metabolites"))

Working_file_tot <- subset(Working_file_tot, !BIO_SAMPLE_ID3 %in% c(""))

data <- tidyr::spread(Working_file_tot, Metabolites, PEAK_AREA_TIC_Log)
data[data == '-Inf'] <- 0
data <- na.omit(data)   


env <- data

library(stringr)
env[c("Date","Numero_sample","ORGANISM", "Time_point","TREATMENT", "Bio.rep")] <- str_split_fixed(env$BIO_SAMPLE_ID3, '_', 6)


env2 <- env[38:43]

row.names(data) <- data$BIO_SAMPLE_ID3
data$BIO_SAMPLE_ID3 <- NULL

library(vegan)
data.dist <- vegdist(data, method="canberra")

# default data by terms

#data.div <- adonis2(data ~ BIO_rep*TREATMENT_1*Time_point*ORGANISM, data = env2, permutations = 999, method="canberra")



#Overall 24h
data.div.Time <- adonis2(data ~ TREATMENT/Time_point, data = env2, permutations = 999, method="canberra")
data.div.TREATMENT <- adonis2(data ~ Time_point/TREATMENT, data = env2, permutations = 999, method="canberra")
data.div.SYNCOM <- adonis2(data ~ TREATMENT/Time_point/ORGANISM, data = env2, permutations = 999, method="canberra")


# R-squared value
R_squared_TIME <- data.div.SYNCOM$R2[2]
R_squared_TREATMENT <- data.div.SYNCOM$R2[1]
R_squared_SYNCOM <- data.div.SYNCOM$R2[3]

# p-value
p.value_TIME <- data.div.SYNCOM$`Pr(>F)`[2]
p.value_TREATMENT <- data.div.SYNCOM$`Pr(>F)`[1]
p.value_SYNCOM <- data.div.SYNCOM$`Pr(>F)`[3]
```

## 2.2 - PcoA metabolic profile (Figure 3.d)

Comparative Metabolic Profiling of Bacteria-Fungi SC (BFSC) for Each Wash.      
PcoA on the all metabolic profile for BFSC through the chambers.         

```{r, echo=FALSE, warning=FALSE, message=FALSE, results="hide"}
#http://www.sthda.com/french/articles/38-methodes-des-composantes-principales-dans-r-guide-pratique/73-acp-analyse-en-composantes-principales-avec-r-l-essentiel/ 
library(dplyr)
Working_file <- Met_Final

Working_file <- subset(Working_file, ORGANISM %in% c("X.BF", "X.X"))
Working_file <- subset(Working_file, !TREATMENT2 %in% c("Peat.A.B"))
Working_file <- subset(Working_file, !BIO_SAMPLE_ID3 %in% c("45258_48_X.X_62h_Peat_B.1", "45268_58_X.BF_62h_Peat.A.BF_B.1", "45268_59_X.X_62h_Peat.A.BF_B.1","45258_48_X.X_62h_Peat_B.4", "45258_47_X.X_62h_Peat_B.4"))



Met_Final_filtered <- Working_file %>% select(c("BIO_SAMPLE_ID3","Metabolites", "PEAK_AREA_TIC_Log")) #Change second criteria according to what you want analyse.

Met_Final_filtered_Spr <- tidyr::spread(Met_Final_filtered, key = Metabolites, value = PEAK_AREA_TIC_Log)

Met_Final_filtered_Spr[is.na(Met_Final_filtered_Spr)] <- 0
Met_Final_filtered_Spr[Met_Final_filtered_Spr == '-Inf'] <- 0

#Met_Final_filtered_Spr <- na.omit(Met_Final_filtered_Spr) 


row.names(Met_Final_filtered_Spr) <- Met_Final_filtered_Spr$BIO_SAMPLE_ID3
Met_Final_filtered_Spr$BIO_SAMPLE_ID3 <- NULL
library("FactoMineR")
library("factoextra")


res.pca <- PCA(Met_Final_filtered_Spr, graph = FALSE)

#Check percent of variation
plot_percent <- fviz_eig(res.pca, addlabels = TRUE, ylim = c(0, 100))

Dim1 <- capture.output(cat(round(res.pca$eig[1,2], digits = 2), "%"))
Dim2 <- capture.output(cat(round(res.pca$eig[2,2], digits = 2), "%"))

res.pca$eig
res.pca$svd

library(data.table)
data_plotting <- data.frame(res.pca$ind$coord)
data_plotting$SP1 <- row.names(data_plotting)

#Concatenate metadata

library(stringr)
data_plotting2 <- data_plotting
data_plotting2[c("Date","Number", "ORGANISM","Time_point", "TREATMENT", "Bio_rep")] <- str_split_fixed(row.names(data_plotting2), '_', 6)

#Remove Pool

data_plotting2 <- subset(data_plotting2, !ORGANISM %in% c("Pool", "pool"))


#Plot
#Claculate for segments
points_pcoa = data.frame(data_plotting2 %>% 
                group_by(Time_point, TREATMENT, ORGANISM) %>% 
                mutate(meanX=mean(Dim.1), meanY=mean(Dim.2)))


#Plot full with all conditions
#Claculate for segments
points_pcoa = data.frame(data_plotting2 %>% 
                group_by(Time_point, TREATMENT, ORGANISM) %>% 
                mutate(meanX=mean(Dim.1), meanY=mean(Dim.2)))
points_pcoa$ORGANISM_TimePoint <- paste(points_pcoa$ORGANISM, points_pcoa$Time_point, sep = "-")

# Create a new column in your data that combines ORGANISM and New_Parameter
data_plotting2$ORGANISM_TimePoint <- paste(data_plotting2$ORGANISM, data_plotting2$Time_point, sep = "-")


# Now, update your ggplot code to include this new parameter
Figure3d <- ggplot() + 
   geom_segment(data = points_pcoa, mapping = aes(x = meanX, y = meanY, xend = Dim.1, yend = Dim.2, color = TREATMENT), size = 0.5, alpha = 0.4) +
  geom_jitter(data = data_plotting2, aes(x = Dim.1, y = Dim.2, shape = ORGANISM_TimePoint, color = TREATMENT), alpha = 1, size = 0.4) + 
  geom_point(data = points_pcoa, aes(x = meanX, y = meanY, shape = ORGANISM_TimePoint, color = TREATMENT), size = 4, alpha = 1) +
  ggtitle("Targeted metabolic profiles (n=41)") + 
   geom_hline(yintercept = 0, linetype = "dashed", color = "#585858") +
  geom_vline(xintercept = 0, linetype = "dashed", color = "#585858") +
  xlab(paste("PCoA1", Dim1)) + ylab(paste("PCoA2",Dim2)) +
  theme(strip.background = element_rect(color = "white", fill = "#585858", size = 1, linetype = "solid"), 
        strip.text.x = element_text(size = 12, color = "white"), 
        strip.text.y = element_text(size = 12, color = "white")) +
  scale_color_manual('Sterile washes', labels = c("Peat", "Peat + Plant", "Peat + Plant + SynCom"), values = c("#DDD5D0", "#988071", "#3B312B")) +
 
  geom_text(aes(x = 2, y = 4.5, label = paste("Time Point : ", round(R_squared_TIME * 100, digits = 3), "% of variance; p =", format(p.value_TIME, digits = 2))), color = "black", size = 2.5) +
  geom_text(aes(x = 2, y = 4, label = paste("Sterile washes : ", round(R_squared_TREATMENT * 100, digits = 3), "% of variance; p =", format(p.value_TREATMENT, digits = 2))), color = "black", size = 2.5) +
geom_text(aes(x = 2, y = 3.5, label = paste("SynCom : ",round(R_squared_SYNCOM*100, digits = 3), "% of variance; p =", format(p.value_SYNCOM, digits = 2))), color = "black", size = 2.5) +
  scale_shape_manual(values = c(17, 19, 2, 1), labels = c("BFSC 24h", "BFSC 62h", "MF 24h", "MF 62h")) 
Figure3d
```

## 2.3 - Log2FC plot (Figure 3.e)  

In this part I compare the abundance of metabolites between sterile extracts when inoculated with BF-SynCom Versus Mock (non-inoculated)      
   
### 2.2.1 - Prepare data
```{r, echo=FALSE, warning=FALSE, message=FALSE, results="hide"}
#Prepare table
# - Full LOG2 plot with significant 

DESeq <- Met_Final

#Keep timepoint 62hours and remove the sterile wash conditioned with Plant and Bacteria
DESeq <- subset(DESeq, Time_point %in% c("62h"))
DESeq <- subset(DESeq, ORGANISM2 %in% c("Media.BacteriaFungi", "Media.Media") & !TREATMENT2 %in% c("Peat.A.B"))

#Add concat column                           
library(dplyr)
DESeq$concat <- paste( DESeq$Metabolites, "-", DESeq$TREATMENT2)
         
                                
DESeq_Soil <-DESeq #subset(DESeq, TREATMENT2 %in% c("Peat.A"))  


#Create what to compare
DESeq_Soil$Met_Treat <- paste(DESeq_Soil$TREATMENT2, "_",DESeq_Soil$Metabolites)


#Control Media
DESeq_Soil_med.med.bac <- subset(DESeq_Soil, ORGANISM2 %in% c("Media.Media"))
Table_Peat_mean_bac <- aggregate(DESeq_Soil_med.med.bac$PEAK_AREA_TIC, list(DESeq_Soil_med.med.bac$concat), FUN=median)
colnames(Table_Peat_mean_bac) <- c("concat", "Mean_Media")

#Split
library(stringr)
Table_Peat_mean_bac[c("Metabolites","TREATMENT22")] <- str_split_fixed(Table_Peat_mean_bac$concat, ' - ', 2)

#Create what to compare
Table_Peat_mean_bac$Met_Treat <- paste(Table_Peat_mean_bac$TREATMENT22, "_", Table_Peat_mean_bac$Metabolites)

#Prepare merge
DESeq_Soil.bac <- subset(DESeq_Soil, ORGANISM2 %in% c("Media.BacteriaFungi"))

#Final merge
DESeq_Soil.bac.f <-  merge(DESeq_Soil.bac, Table_Peat_mean_bac,  by = c("Met_Treat"))  

#Calculate log2FC
DESeq_Soil.bac.f['Mean_Media'][DESeq_Soil.bac.f['Mean_Media'] == 0] <- 0.1

DESeq_Soil.bac.f$log2FC <- (log2(DESeq_Soil.bac.f$PEAK_AREA_TIC))-(log2(DESeq_Soil.bac.f$Mean_Media))


#_INF = Not detected put -5
DESeq_Soil.bac.f
DESeq_Soil.bac.f['log2FC'][DESeq_Soil.bac.f['log2FC'] == '-Inf'] <- -5
DESeq_Soil.bac.f['log2FC'][DESeq_Soil.bac.f['log2FC'] <  -5] <- -5

#DATA_Log2_PEAT2 <-  subset(DATA_Log2_PEAT2, !log2FC %in% c("NaN"))

#DESeq_Soil.bac.f['log2FC'][DESeq_Soil.bac.f['log2FC'] == '0'] <- -5
#DESeq_Soil.bac.f['log2FC'][DESeq_Soil.bac.f['log2FC'] == '0'] <- -5


#Prepare color
DESeq_Soil.bac.f$TREATMENT22

Table_S <- DESeq_Soil.bac.f %>% group_by(ORGANISM2,Metabolites.y,TREATMENT22) %>% summarise (log2FC_med = median(log2FC))

Table_S$concat <-  paste(Table_S$Metabolites.y, "-",Table_S$ORGANISM2, "_", Table_S$TREATMENT22 )

Table_S$ORGANISM2 <- NULL
Table_S$Metabolites.y <- NULL
Table_S$TREATMENT22 <- NULL

DESeq_Soil.bac.f$concat <-  paste(DESeq_Soil.bac.f$Metabolites.y, "-",DESeq_Soil.bac.f$ORGANISM2, "_", DESeq_Soil.bac.f$TREATMENT22 )



Final_tab <- merge(x = Table_S, y = DESeq_Soil.bac.f, by = c("concat"))

#Keep only what is significant
Final_tab$Significant <-   ifelse(Final_tab$log2FC_med >= 1 , "Significant",
                           ifelse(Final_tab$log2FC_med <= -1, "Significant","Non Significant"))

DATA_Log2_PEAT2_f <- Final_tab
```

### 2.2.2 - Prepare data

**Stats:**    
We performed a Kruskal-Wallis test followed by a post-hoc Dunn test to assess the significance of differences in Metabolites abundance between Microbial conditions (BSC, FSC and BFSC) for each specific wash.      

```{r, echo=FALSE, warning=FALSE, message=FALSE, results="hide"}

#Calculate significance between BSC, FSC and BFSC for each individual metabolite in each individual wash.
library(dplyr)
library(dunn.test)

#Load dataset Change accordingle (Table_f_S or Table_f_C)
df <- Final_tab

#Filter one with -7
df_filtered <- subset(df, !Metabolites.x %in% c("alpha.ketoglutaric.acid", "fumaric.acid") )

#Load again
df <- df_filtered
df$Metabolites.x_Class <- paste0(df$Metabolites.x, "/", df$Class)


# Get unique conditions
unique_conditions <- unique(df$Metabolites.x_Class)

# Create empty data frames to store the results
kruskal_results <- data.frame()
dunn_results <- data.frame()

# Loop over unique conditions and perform tests
for (condition in unique_conditions) {
  # Subset the data for the current condition
  subset_data <- df[df$Metabolites.x_Class == condition, ]
  
  # Perform Kruskal-Wallis test
  kruskal_result <- kruskal.test(log2FC ~ TREATMENT2, data = subset_data)
  
  # Perform Dunn's test for post hoc comparisons
  dunn_result <- dunn.test(subset_data$log2FC, g = subset_data$TREATMENT2, method = "bh")
  
  # Store the results in data frames
  kruskal_results <- rbind(kruskal_results, c(Condition = condition, kruskal_result$P.adjusted))
  dunn_results <- rbind(dunn_results, data.frame(dunn_result$comparison, dunn_result$P.adjusted, condition))
}

# Print Kruskal-Wallis results
print(kruskal_results)

# Print Dunn test results
print(dunn_results)

#Create groups letters
library(rcompanion)

# Create empty data frames to store the results
groups_results <- data.frame()

for (condition in unique_conditions) {
  # Subset the data for the current condition
  subset_data <- dunn_results[dunn_results$condition == condition, ]
   
  group <- cldList(comparison = subset_data$dunn_result.comparison, 
                    p.value    = subset_data$dunn_result.P.adjusted,
                    threshold  = 0.05)
  # Create a data frame for the results of the current condition
  condition_results <- data.frame(Condition = condition,
                                  Group = group$Group,
                                  Letter = group$Letter,
                                  MonoLetter = group$MonoLetter)
  
# Append the results to the groups_results data frame
  groups_results <- rbind(groups_results, condition_results)

}
groups_results

# Splitting the column into two
library(tidyr)
groups_results <- separate(groups_results, Condition, into = c("Metabolite.y", "Class"), sep = "/")

#Arrange table to match the plot
colnames(groups_results) <- c("Metabolites.y","Class","TREATMENT2", "Letter", "MonoLetter")
groups_results$ORGANISM2 <- c("Media.BacteriaFungi")
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results="hide"}

#REmove Metabolites which are not signficiant in any condition
Final_tab$remove <- ifelse(Final_tab$Significant == "Non Significant", 1,2)

REMOVE <- Final_tab %>%  
  group_by(Metabolites.y) %>% 
  summarize(mean=mean(remove)) 

To_remove <- subset(REMOVE, mean %in% c(1) )


#Remove eveything between 0.5 and -0.5 (non significant)
Final_tab_f <- subset(Final_tab, !Metabolites.y %in% c(To_remove$Metabolites.y))
groups_results <- subset(groups_results, !Metabolites.y %in% c(To_remove$Metabolites.y))

#Change 3PG 2PG
Final_tab_f$Metabolites.y <- ifelse(Final_tab_f$Metabolites.y == "X2.phosphoglyceric.acid..2PG..and..3PG.", "2PG.and.3PG",Final_tab_f$Metabolites.y) 


groups_results$Metabolites.y <- ifelse(groups_results$Metabolites.y == "X2.phosphoglyceric.acid..2PG..and..3PG.", "2PG.and.3PG",groups_results$Metabolites.y) 

#Remove GABA (non significant)
Final_tab_f <- subset(Final_tab_f, !Metabolites.y %in% c("GABA_1Bz"))
groups_results <- subset(groups_results, !Metabolites.y %in% c("GABA_1Bz"))

#Plot
#Order
x = tapply(Final_tab_f$log2FC_med, Final_tab_f$Metabolites.y, function(x) max(x))
x = sort(x, FALSE)
Final_tab_f$Metabolites.y = factor(as.character(Final_tab_f$Metabolites.y), levels=names(x))

#PLOT

#Order
x = tapply(Final_tab_f$log2FC_med, Final_tab_f$Metabolites.y, function(x) max(x))
x = sort(x, FALSE)
Final_tab_f$Metabolites.y = factor(as.character(Final_tab_f$Metabolites.y), levels=names(x))


#Change facet_grid names
Class_names <- setNames(c("Amino Acids", "TCA Glycolysis"), c("Amino_Acids", "TCA_Glycolysis"))


#Plot
Plot_bac <- ggplot(data=Final_tab_f, 
                             aes_string(x='Metabolites.y',y='log2FC', fill = 'TREATMENT2')) + # Significant Or ORGANISM2
  geom_boxplot(outlier.shape = NA, alpha = 0.8, size = 0.2) +
  geom_point(alpha=1, size = 0.5, position = position_dodge(width=0.75),data=Final_tab_f, 
                             aes_string(color = 'TREATMENT2')) +
  theme(
        strip.background = element_rect(color="white", fill="#585858", size=1, linetype="solid"),
        strip.text.x = element_text(size = 12, color = "white"),
        strip.text.y = element_text(size = 12, color = "white"),
        axis.title.y = element_text( size = 12),
        plot.title = element_text(size = 16),
        axis.text.x = element_text(angle = 90,  hjust=1), 
        axis.text.y = element_text())+  
  ylab ("Log2FC") + xlab ("") +
   facet_grid(~Class, space="free", scale = "free", labeller = labeller(Class =Class_names)) +
  geom_hline(yintercept = c(0.5,-0.5), linetype="dashed", size=0.5) +
  ggtitle("Log 2 FC relative to the washes")+
       scale_fill_manual('Sterile washes',labels = c("Peat", "Peat + Plant", "Peat + Plant + SynCom"), values=c("#DDD5D0", "#988071", "#3B312B" )) +
     scale_color_manual('Sterile washes',labels = c("Peat", "Peat + Plant", "Peat + Plant + SynCom"), values=c("#DDD5D0", "#988071", "#3B312B" )) +
        geom_text( data    = groups_results, mapping = aes(x =Metabolites.y , y = 4, label = Letter), size = 4,position = position_dodge(width=0.75))
Plot_bac #CE PLOT



```

*****

# End of script
      
Author : Guillaume Chesneau     
[X](https://twitter.com/Chesneau_G).   
[Researchgate](https://www.researchgate.net/profile/Guillaume-Chesneau-3).   
[Google Scholar](https://scholar.google.fr/citations?user=ywzDf3QAAAAJ&hl=fr&oi=ao). 

MPIPZ Cologne,  [Hacquard's group](https://www.mpipz.mpg.de/hacquard).   

*****